---
title: "episode-2"
format: html
---

```{r}
library(tidyverse)
library(terra)
```
Loading the raster

```{r}
DSM_HARV <- rast("data/NEON-DS-Airborne-Remote-Sensing/HARV/DSM/HARV_dsmCrop.tif")

DSM_HARV
```

Transform raster. into a dataframe

```{r}
DSM_HARV_df <- as.data.frame(DSM_HARV, xy = TRUE)
```

Quick plot to check things

```{r}
ggplot() +
    geom_raster(data = DSM_HARV_df , aes(x = x, y = y, fill = HARV_dsmCrop)) +
    scale_fill_viridis_c() +
    coord_quickmap()
```
create the column with the binned categories

```{r}
DSM_HARV_df <- DSM_HARV_df %>%
                mutate(fct_elevation = cut(HARV_dsmCrop, breaks = 3))
```

Let's check this

```{r}
ggplot() +
    geom_bar(data = DSM_HARV_df, aes(fct_elevation))
```

Check the distribution of those bins:

```{r}
unique(DSM_HARV_df$fct_elevation)
```

```{r}
DSM_HARV_df %>%
        group_by(fct_elevation) %>%
        count()
```

Those numbers might not be the best to use, so we might what to customize those threshold values


```{r}
custom_bins <- c(300, 350, 400, 450)

DSM_HARV_df <- DSM_HARV_df %>%
  mutate(fct_elevation_2 = cut(HARV_dsmCrop, breaks = custom_bins))

unique(DSM_HARV_df$fct_elevation_2)
```

Let's have a look at the distribution

```{r}
ggplot() +
  geom_bar(data = DSM_HARV_df, aes(fct_elevation_2))
```

```{r}
DSM_HARV_df %>%
  count(fct_elevation_2) 
```

Ok let's use this in a plot:

```{r}
ggplot() +
  geom_raster(data = DSM_HARV_df , aes(x = x, y = y, fill = fct_elevation_2)) + 
  coord_quickmap()
```

OK the color are not optimal for elevation

```{r}
ggplot() +
 geom_raster(data = DSM_HARV_df , aes(x = x, y = y,
                                      fill = fct_elevation_2)) + 
    scale_fill_manual(values = terrain.colors(3)) + 
    coord_quickmap()
```

OK let's keep improving the formatting


```{r}
my_col <- terrain.colors(3)

ggplot() +
 geom_raster(data = DSM_HARV_df , aes(x = x, y = y,
                                      fill = fct_elevation_2)) + 
  scale_fill_manual(values = my_col, name = "Elevation") + 
  theme(axis.title = element_blank()) + 
  coord_quickmap()
```

## Layering rasters

DSM hillshade

```{r}
DSM_hill_HARV <-
  rast("data/NEON-DS-Airborne-Remote-Sensing/HARV/DSM/HARV_DSMhill.tif")

DSM_hill_HARV
```

```{r}
DSM_hill_HARV_df <- as.data.frame(DSM_hill_HARV, xy = TRUE) 

str(DSM_hill_HARV_df)
```

```{r}
ggplot() +
  geom_raster(data = DSM_hill_HARV_df,
              aes(x = x, y = y, alpha = HARV_DSMhill)) + 
  scale_alpha(range =  c(0.15, 0.65)) +  
  coord_quickmap()
```


```{r}
ggplot() +
  geom_raster(data = DSM_hill_HARV_df,
              aes(x = x, y = y,
                  alpha = HARV_DSMhill)) +
  geom_raster(data = DSM_HARV_df ,
              aes(x = x, y = y,
                  fill = HARV_dsmCrop,
                  alpha = 0.6)) +
  scale_fill_viridis_c() +  
  scale_alpha(range = c(0.15, 0.65), guide = "none") +  
  ggtitle("Elevation with hillshade") +
  coord_quickmap()
```
